Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-02-09 SIR_manuscript.tex
edited start of the discussion

Mathematical modeling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation. 
Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics. 
While targeted testing strategies 
\david{you need to be clearer: you mean targeting individuals who have contacted people who were infected.  For example, here you could say something like ``While targeting the contacts of confirmed cases\dots''}
are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread
is greater for a slow test than for a fast test. This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).

We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies. Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $W_I/W_S=3$ and $W_R=W_I$. Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario. 
This part needs to be developed further in future work. 
Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment. 
This can be implied when we would like to quantify and compare the effect of different levels of test forcus \david{??} for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management. For example, when people are tested for "screening", the individuals with potential higher mobility, eg. people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with. With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing processes.

>>>>

Mathematical modelling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation. Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time, and isolation on epidemic dynamics.

In agreement with our intuition, targeted testing strategies (i.e., strategies that identify the most infectious cases) are always more effective than random testing at lowering $\Rnum$. This reinforces the benefit of contact tracing in slowing transmission since tracing contacts of positive cases increases the likelihood of identifying other positive cases. Another approach that may enhance targeting accuracy may be to use pre-testing selection criteria that must be met by individuals wishing to be tested (e.g., symptoms questionnaires). While such an approach may enhance targeting, it may carry significant downsides (e.g., lessening the surveillance breadth of testing), which must be judged against the benefits.
\david{you need to be clearer: you mean targeting individuals who have contacted people who were infected.  For example, here you could say something like ``While targeting the contacts of confirmed cases\dots''}\fady{I moved this comment below the paragraph (so David is referring to the previous paragraph) and tried to resolve it. What do you think?}

Less intuitively, our analysis suggests that there are cases where reporting test results more rapidly could increase, rather than decrease $\Rnum$. This can occur when individuals are much more cautious when awaiting a test result than when they believe they are not infected, and may not be robust to second-order effects of fast testing (such as better contact tracing).

We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies. In the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals (thus $W_I/W_S=3$ and $W_R=W_I$). Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.\fady{I don't understand the previous sentence}
This part needs to be developed further in future work. 
Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment. 
This can be implied when we would like to quantify and compare the effect of different levels of test forcus \david{??}\fady{I don't get this part either.} for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management. For example, when people are tested for "screening", the individuals with potential higher mobility, eg. people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with. With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing processes.
